Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2020-05-14 article.tex
Assume immunodominance, instead of having to prove the shape of the fitness landscape

\paragraph{Test that selection is additive}

This study assumes a gradual fitness landscape, in which it always
(slightly) beneficial for a pathogen to lose an epitope.
Because this assumption needs to be correct for the research
to be viable, we did an
exploratory experiment (see 'Fitness landscape', 
subsection \ref{subsec:fitness_landscape}).
We concluded that the assumption may be correct 
\richel{Let's assume for now}.

@@

We thank the Center for Information Technology of the University 
of Groningen for its support and for providing access to the Peregrine 
high performance computing cluster.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Fitness landscape}
\label{subsec:fitness_landscape}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\paragraph{Introduction}
One could assume a two-state fitness landscape: being
undetected, due to zero epitopes, or being detected, due to one or more
epitopes. Alternatively, one may hypothesize a gradual fitness decrease per
extra epitope. Both fitness landscapes are depicted in 
figure \ref{fig:fitness_landscape}.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fitness_landscape.png}
  \caption{
    Two conceptual fitness landscapes. At the left is a two-state
    fitness landscape, in which the number of epitopes has a simple
    effect: a pathogen's fitness is high if it has no epitopes,
    as it is undetected by a host's immune system. 
    For one or more epitopes, fitness is at a same lower values. 
    In this fitness landscape, acquiring a second or third or more 
    epitope, this has no effect on fitness.
    At the right is a gradual fitness landscape, in which any increase in
    the number of epitopes results in a lower fitness.
  }
  \label{fig:fitness_landscape}
\end{figure}

\paragraph{Why}
This exploratory experiment tests an assumptions 
of the bigger evolutionary experiment. 
If the assumption is incorrect,
a different evolutionary experiment needs to be contrived.

\paragraph{Method}
For a given MHC-I haplotype, we count the number of epitopes in 
the proteome of COVID-19, at the start of the 2019 pandemic (as
described in \cite{wu2020new})
and a random strain sequenced most recently at the day of the experiment. 
Of both proteomes, we use use \verb;epitope-prediction; \cite{bianchi2017} 
to count the number of predicted binders for the HLA-A-02:01 haplotype.
We use that haplotype, as it is the haplotype used by default with the
\verb;epitope-prediction; package.

\paragraph{Decision rule}
If the number of epitopes has increased, we assume a two-state fitness landscape,
in which the number of epitopes can increase neutrally by drift alone.
In this case, the experiment will be cancelled and replaced by a 
viable approach.
Else, the bigger evolutionary experiment is done as described already.
It may still be possible that the number of epitopes decreased due to drift,
but this would be a proper result in confirming the null hypothesis
of $\mathcal{H}_{3,HVB}$.

\paragraph{Method detail}
Because the proteome is only sequenced at the start of the pandemic,
we use the DNA sequences of both strains, and infer the proteome
of each, using Prodigal. Although the earlier strand has a published
proteome, we infer it regardless, as to use the same pipeline.
If needed, we tweak the Prodigal parameter settings up until
the inferred proteome of the earlier strain with known proteome
matches well enough, as concluded by eyeballing.
The parameter settings used to infer a good proteome on the earlier
strain, is then used on the genome of the more recent strain 
to infer our recent proteome.

>>>>

\paragraph{Immunodominance}

The effective number of epitopes, however,
is much lower than the actual number of possible
epitopes, due to immunodominance. 
Immunodominance is a feature of the immune system,
due to which some epitopes dominate in causing an immune response,
where other (called subdominant) epitopes end up having no 
effect \cite{akram2012immunodominance}.
The same study mentions 20 factors that influence the strength
of an immune response.
The most relevant factor for this study is the effect of
a mutation within an epitope, 
such as the in vitro example described in \cite{berkhoff2004mutation},
where one single mutation resulted in a 5-20\% weaker
responese of virus-specific CD8+ T cells \richel{Fun experiment:
do this in silico}.
Due to immunodominance, we dare assume that there can be selection
for immune system avoidance, as the effective number of loci may be
low and effects of a single mutation noticable.

@@

We thank the Center for Information Technology of the University 
of Groningen for its support and for providing access to the Peregrine 
high performance computing cluster. 
Additionally, we would like to thank Sci-Hub \cite{himmelstein2018sci}
for allowing us to read paywalled articles while working from home.

@@
